---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(project)
```

# Conduct a Statistical Test on the Human Population

<!-- badges: start -->
[![R-CMD-check](https://github.com/MQ-STAT1378/project-nzh-zhou/workflows/R-CMD-check/badge.svg)](https://github.com/MQ-STAT1378/project-nzh-zhou/actions)
<!-- badges: end -->

This package contains a set of functions to test one of three questions about the human population. Read section 1 of the following document for more information about the three questions.

``` {r report, eval = FALSE}
file.edit('vignettes/articles/report.Rmd')
```

## Installation

You can install this package via github. Run the following command.

``` {r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("https://github.com/MQ-STAT1378/project-nzh-zhou", 
                         auth_token = "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")
```

## Functions and Data

This package contains the data set, `pro`, which can be used as an example for the functions included in this package.

There are 5 helper functions, wherein each one outputs part of the statistical test for a specific question, and they all take the same three inputs.

* `data`: the data input; must be of class "data.frame"
* `question`: the question number; must be an integer between 1 and 3 inclusive
* `output`: the desired type of output; must be either "LaTeX" or "R"

The wrapper function, `mytest`, outputs the whole statistical test for a specific question in R format, but only takes in inputs `data` and `question`.

## Examples

```{r examples}
```
